---
title: "Day 3"
author: "Lucas Anderton"
date: "1/4/2020"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Statistics with R

### The Root-Mean-Square (RMS)
RMS describes average *magnitude* of variable's values. The RMS takes each values,

1. squares it,
2. takes the mean of these squares, and then
3. takes the square root.

Why take the square, then square root?

### Standard Deviation (SD)
SD describes the spread of a variable.
SD: the RMS of the deviations from the average. Tells us "How far from the average is a typical value of the variable?" Calculate by taking each observation's difference from the average, then take the RMS of those differences. Use `sd` to calculate in R.  

### Variance (SD^2^)
Variance: SD squared.
Variance: average of the squared deviations from the mean. Use `var` to calculate it in R.

### The z-score
For variable X, the z-score of observation xi, tells how far it is from average, in units of the standard deviation. 

Inteperetation: z-score does not depend on units we measure in (as long as linnear transformation). So, for example, we could compare currencies.

#### Practice with two respondents.
```{r}
resp1_inc <- (65000 - 50000) / 15000
resp1_inc
resp2_inc <- (20000 - 50000) / 15000
resp2_inc
```

### QQ plot
Visually compare the quantiles of two distributions.

## Probability
Foundations:

1. **Experiment**: a process that yields a probailistic/stochastic outcome. Not nec. entirely random, but with soem random component.
2. **Outcome space/Sample space**: The set of all possible outcomes of an experiment. Usually denoted by Omega $\Omega$.
3. **Event**: A subset of Omega. Usually denoted A, B, etc. The probability of A happening is P(A).
4. **Complement**:

Examples:

1. Experiment: a voter will vote Dem, vote Rep, vot other, or abstain.
2. Outcome space: Omega $\Omega$ = {Dem, Rep, other, abstainn}
3. Event: 
    - A = abstains. Assuming all equally likely, what is P(A)?
    - B = supports a major party candidate. Assuminng all equally likely, what is P(B)?
4. Complement:
    - A^C^ is the probability they vote for a Democrat, Republican, or third party candidate, so 75%.
    - B^C^ is the probability they don't vote or don't vote for a major party. 

### Probability of either of 2 events
$P(A$ or $B) = P(A) + P(B) - P(A$ and $B)$

### Law of total probability
We can decompose probability of $A$ into two components: $A$ happening when $B$ also happens, and $A$ happening when "not $B$" happens. 

### Combinations: Counting selected sets
How many ways to **select** $k$ things from a set of $n$ things?

### Indpendence
Events are **independent** if they don't provide any innformation about each other. Knnowinng A happened doesn't change the probability of B happening. Know B happening doesn't change the probability of A happening. 

If A and B are independent, then:   
  P(A and B) = P(A)P(B|A)  
            = P(A)P(B)  

## A probability example
For a given Conngress,

1. Take a random sample of 10 bills.
2. Calculate p, the prop in your sample that passed. 
3. Calculate the Standard Error around the p
4. Find the critical value for an 80% interval
5. Calculate an 80% confidence innterval aroundn p

```{r}
samp <- rbinom(10, 1, .35)
samp

phat <- mean(samp)
se <- sqrt(phat * (1 - phat) / 10)
phat
se

critval <- qnorm(.9)
lower <- phat - critval * se 
upper <- phat + critval * se
lower
upper

```







  