---
title: "Day 4"
author: "Lucas Anderton"
date: "1/6/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Running Lowess

```{r}
x <- seq(1,25,length = 600)
y <- (2/(pi*x))^(0.5)*(1-cos(x)) + rnorm(100,0,1/10)
summary(lm(y~x))$coef

#postscript("Class.Multilevel/trends1.ps")
par(mar = c(3, 3, 2, 2), bg = "white") # adjusts the marginns and background
plot(x, y, pch = "+") # sets the plot points as a plus
ols.object <- lm(y ~ x)
abline(ols.object, col = "blue") # plots the regression line
lo.object <- lowess(y ~ x, f = 2/3) # smooths the line; the bigger the number, the smoother it is
lines(lo.object$x, lo.object$y, lwd = 2, col = "red")
lo.object <- lowess(y ~ x, f = 1/5)
lines(lo.object$x, lo.object$y, lwd = 2, col = "purple")
#dev.off()
```

## What does the Linear Model get you?

```{r}
anaemia <- read.table("http://jeffgill.org/files/jeffgill/files/anaemia.dat_.txt", header=TRUE, row.names=1)
a.lm.out <- lm(Hb ~ Age + PCV, data=anaemia)
summary(a.lm.out)
```

Check out `cooks.distance` for determining how influential a value is on regression. 

## Milticollinearity: Remedies

- Respecify model (if reasonable): add/drop variables, add data cases that break the patternn, restrict the rannge of some variables, combine variables possibly with PCA. 
- Center explanatory variables, or stanndardize (slope coefficient is interpreted in nunnits of standard deviations of the covariate, the interecept is the mean of the outcome `y` when all covariate values are zero).
- Create a new variable that is a weighted combination of highly correlated variables and use it to replace both (two variables to one variable in the model)
- Ridge regression (see below)

## Simple Ridge Regression Example

```{r}
anaemia <- read.table("http://jeffgill.org/files/jeffgill/files/anaemia.dat_.txt", header=TRUE, row.names=1)
library(MASS)
a.lm3.out <- lm.ridge(Hb ~ Age + Menapouse + I(Age + rnorm(nrow(anaemia))), data = anaemia)
a.lm3.out$GCV

cbind(a.lm3.out$coef, sqrt(a.lm3.out$scales))

summary(lm(Hb ~ Age + Menapouse, data = anaemia))$coef[2:3,1:2]
```



